---
title: 'Reported Statistics Appendix'
output: 
    html_document:
        toc: true
        toc_float: yes
        code_folding: hide
---

```{r setup, include=FALSE}
library(lme4)
library(lmerTest)
library(lmtest)
library(sandwich)  # For clustered SEs
library(stargazer)
library(effects)
library(tidyverse)

theme_set(theme_minimal())

# Loading data
all_complete_tables <- str_subset(dir('Output/Tables/'), 'complete')
most_recent <- max(str_extract(all_complete_tables, '\\d{4}-\\d{2}-\\d{2}'))

complete_table <- read_delim(str_c('Output//Tables//',
                                 str_subset(all_complete_tables, most_recent))[1], delim = ';')

dat_main_long <- read_delim('..//Data//Clean//all_participants_long_main_main_study.csv',   
                            delim = ';')

# Color secifications for the plots
col_set <- c('lightgrey', 'grey', 'darkgrey', 'black')
```

<!-- #################################################################### -->
# Referenced in Phase 1

## Inverse Updates
```{r inverse_updates_gender}
dat_prepared <- dat_main_long %>%
  filter(i_block < 2) %>%
  group_by(participant_code, gender) %>%
  summarize(n_wrongs = sum(updating_type == 'Wrong'))

wilcox.test(n_wrongs ~ gender, data = dat_prepared)
```
No more inverse updates for either gender.

Correlations:
Replace the variable in `cor.test()` with any of the ones in the `group_by()` statement to test their correlations.
```{r inverse_updates_corrs}
dat_prepared <- dat_main_long %>%
  filter( #age < 40,
    i_block < 2) %>%
  group_by(participant_code, investment_experience,
    age, engagement, cogn_rpm_total_points, risk_soep_general) %>%
  summarize(n_wrongs = sum(updating_type == 'Wrong'))

cor.test(dat_prepared$n_wrongs, dat_prepared$risk_soep_general,
  method = 'spearman')

```
```{r inverse_updates_de}
dat_prepared <- dat_main_long %>%
  filter(i_block < 2) %>%
  group_by(participant_code) %>%
  summarize(n_wrongs = sum(updating_type == 'Wrong')) %>%
  left_join(complete_table, by = 'participant_code')

cor.test(dat_prepared$n_wrongs, dat_prepared$de_12)
```

<!-- -------------------------------------------------------------------- -->
## The Role of Beliefs

Did higher engagement lead to a "more reasonable" scale in belief reports?
```{r belief_var_by_engagement}
dat_prepared <- dat_main_long %>%
  dplyr::filter(i_block < 2) %>%
  group_by(participant_code, engagement) %>%
  summarize(belief_sd = sd(belief, na.rm = TRUE))

cor.test(dat_prepared$engagement, dat_prepared$belief_sd,
  method = 'spearman')
```

Variance of the belief reports between subjects.
```{r belief_variance_hist, message = FALSE}
these_sds <- dat_main_long  %>%
  filter(i_block < 2) %>%
  group_by(participant_code) %>%
  summarise(sd_per_participant = sd(belief / 100, na.rm = TRUE))

ggplot(these_sds, aes(x = sd_per_participant)) +
  geom_histogram(alpha = .6, fill = 'black') +
  geom_vline(xintercept = .03135, linetype = 'dashed') +
  labs(x = 'Belief Report Standard Deviations', y = 'Frequency')

# ggsave('belief_variance_hist.pdf',  device = 'pdf',
#   width = 10, height = 7, path = file.path('Output', 'Plots'))
```



<!-- -------------------------------------------------------------------- -->
### Belief Formation
Frequency of update types compared to Bayesian updating.
```{r update_type_table}
knitr::kable(prop.table(table(dat_main_long$updating_type)),
  col.names = c('Updating Type', 'Proportion'))
```

How strong is the updating after `n` up or down signals?
```{r updating_magnitude_by_run_length}
dat_prepared <- dat_main_long %>%
  dplyr::filter(i_block < 2) %>%
  mutate(run_position = unlist(sapply(rle(price_move_from_last)$lengths,
    seq_len))) %>%
  group_by(price_move_from_last, run_position) %>%
  summarize(
    avg_update = mean(belief_diff_from_last / 100, na.rm = TRUE),
    SE = sd(belief_diff_from_last / 100, na.rm = TRUE) / sqrt(n()),
    lower_CI = avg_update + qt(.05, n()) * SE,
    upper_CI = avg_update + qt(.95, n()) * SE) %>%
  filter(run_position < 7,
    !is.na(price_move_from_last))

ggplot(dat_prepared,
  aes(x = run_position, y = avg_update,
    group = price_move_from_last, color = price_move_from_last)) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), width = .05) +
  scale_color_manual(values = c(col_set[1], col_set[3])) +
  labs(x = 'N Same Moves', y = 'Mean Update', color = 'Price Move',
    group = 'Price Move')

# ggsave('updating_magnitude_by_run_length.pdf', device = 'pdf',
#   width = 10, height = 7, path = file.path('Output', 'Plots'))
```

The same but with the beliefs
```{r beliefs_by_run_length}
dat_prepared <- dat_main_long %>%
  dplyr::filter(i_block < 2) %>%
  mutate(run_position = unlist(sapply(rle(price_move_from_last)$lengths,
    seq_len))) %>%
  group_by(price_move_from_last, run_position) %>%
  summarize(
    avg_belief = mean(belief / 100, na.rm = TRUE),
    SE = sd(belief / 100, na.rm = TRUE) / sqrt(n()),
    lower_CI = avg_belief + qt(.05, n()) * SE,
    upper_CI = avg_belief + qt(.95, n()) * SE) %>%
  dplyr::filter(run_position < 7,
    !is.na(price_move_from_last))

ggplot(dat_prepared,
  aes(x = run_position, y = avg_belief,
    group = price_move_from_last, color = price_move_from_last)) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), width = .05) +
  scale_color_manual(values = c(col_set[1], col_set[3])) +
  labs(x = 'N Same Moves', y = 'Mean Belief', color = 'Price Move',
    group = 'Price Move')

# ggsave('beliefs_by_run_length.pdf', device = 'pdf',
#   width = 10, height = 7, path = file.path('Output', 'Plots'))
```

Testing whether the interaction pattern also holds when only looking at long or short investments seperately.
Change the filter `lag(hold) == 1` between `1` and `-1` to either only include long or short investments.
```{r updating_baseline_test_long_short}
dat_prepared <- dat_main_long %>%
  filter(i_block < 2,
         position_end_of_last_period %in% c('Gain', 'Loss'),
         dplyr::lag(hold) == 1,
    ) %>%
  mutate(price_move_from_last_corrected = dplyr::recode(
      price_move_from_last_corrected, 'Unfavorable' = -.5, 'Favorable' = .5),
    position_end_of_last_period = dplyr::recode(
      position_end_of_last_period, 'Loss' = -.5, 'Gain' = .5),
    belief_diff_bayes_corrected_flipped = belief_diff_bayes_corrected_flipped / 100,
    participant_code = participant_code)

belief_updates_baseline <- lm(belief_diff_bayes_corrected_flipped ~
  position_end_of_last_period * price_move_from_last_corrected + age + gender,
   data = dat_prepared)

coeftest(belief_updates_baseline,
         vcovCL(belief_updates_baseline,
           cluster = ~dat_prepared$participant_code))
```

Checking whether there is an interaction in the updating during the first two blocks when correcting for a RL learner. These effects are __dummy coded!__ and clustered by participant. Comment out the `updating_type != 'Wrong'` filter to get the results with wrong updates included.
```{r updating_baseline_test_rl}
dat_prepared <- dat_main_long %>%
  dplyr::filter(i_block < 2,
         position_end_of_last_period %in% c('Gain', 'Loss'),
         updating_type != 'Wrong'
    ) %>%
  mutate(price_move_from_last_corrected = recode(
      price_move_from_last_corrected, 'Unfavorable' = 0, 'Favorable' = 1),
    position_end_of_last_period = recode(position_end_of_last_period,
      'Loss' = 0, 'Gain' = 1))

belief_updates_baseline <- lm(belief_diff_rl_corrected_flipped ~
  position_end_of_last_period * price_move_from_last_corrected,
   data = dat_prepared)

coeftest(belief_updates_baseline,
         vcovCL(belief_updates_baseline,
           cluster = ~dat_prepared$participant_code))
```
_(robust to excluding inattentives)_

Boxplots about the variances of PGR and PLR.
To get the plot using the raw values (not relative to the rational benchmark) comment out the recoding line for `Value` in the `mutate()` call.
```{r pgr_plr_phase_one}
complete_table %>%
  dplyr::select(participant_code, plr_12, pgr_12) %>%
  pivot_longer(-participant_code, names_to = 'Position',
  values_to = 'Value') %>%
  mutate(Position = dplyr::recode(
    Position, 'pgr_12' = 'PGR', 'plr_12' = 'PLR'),
    Value = if_else(Position == 'PGR', Value - .19, Value - .97)
    ) %>%

ggplot(aes(x = Position, y = Value)) +
  geom_line(aes(group = participant_code), alpha = .25,
    position = position_jitter(.05)) +
  geom_boxplot(alpha = .5, outlier.shape = NA)

# ggsave('pgr_plr_phase_one_boxplots_benchmark.pdf', device = 'pdf',
#   width = 10, height = 7, path = file.path('Output', 'Plots'))
```


<!-- #################################################################### -->
# Referenced in Phase 2

## Trading Behavior

DE of the first phase corrected for the benchmark. There is no significant difference between groups in the first phase as all participants received the same treatment.
```{r de_conditions_baseline}
summary(lm(de_12 + .78 ~ main_condition + age + gender, data = complete_table))
```

Finding out whether participants adhered to the shown probabilities in the states_shown condition.
```{r beliefs_states_shown}
dat_main_long %>%
  dplyr::filter(condition == 'states_shown') %>%
  mutate(state = factor(state, levels = 0:1, labels = c('Bad', 'Good'))) %>%
  group_by(participant_code, state) %>%
  summarize(mean_belief = mean(belief) / 100,
            sd_belief = sd(belief / 100)) %>%
  mutate(diff_to_rational = if_else(state == 'Good', .65 - mean_belief,
            .35 - mean_belief)) %>%
  group_by(participant_code) %>%
  mutate(rationality_score = sum(abs(diff_to_rational))) %>%
  arrange(rationality_score) %>%
  ungroup() %>%
  mutate(participant_num = factor(
    participant_code, levels = unique(participant_code))) %>%

  ggplot(aes(x = participant_num, y = mean_belief, color = state)) +
  geom_point(position = position_dodge(.25)) +
  geom_errorbar(aes(ymin = mean_belief - sd_belief,
    ymax = mean_belief + sd_belief, width = 0),
    position = position_dodge(.25)) +
  scale_color_manual(values = c('black', 'darkgrey')) +
  labs(x = 'Participant', y = 'Average Belief', color = 'State') +
  theme(axis.text.x = element_blank())

# ggsave('avg_beliefs_states_shown_bw.pdf',  device = 'pdf',
#   width = 10, height = 7, path = file.path('Output', 'Plots'))
```


<!-- #################################################################### -->
## Belief Updating

```{r updating_states_shown_hist}
ggplot(dplyr::filter(dat_main_long,
  position_end_of_last_period %in% c('Gain', 'Loss'),
         condition == 'states_shown', i_block > 1),
  aes(x = belief_diff_from_last / 100)) +
  facet_grid(rows = vars(position_end_of_last_period,
    price_move_from_last_corrected)) +
  geom_histogram(fill = col_set[3]) +
  labs(x = 'Belief Updates in full information Condition', y = 'Frequency')

ggsave('updating_full_info_hist.pdf',  device = 'pdf',
  width = 10, height = 7, path = file.path('Output', 'Plots'))
```

```{r updating_states_shown_bars}
dat_prepared <- dat_main_long %>%
  dplyr::filter(position_end_of_last_period != 'No Returns',
         condition == 'states_shown',
         i_block > 1) %>%
  mutate(belief_diff_from_last =
    belief_diff_bayes_corrected_flipped / 100)

dat_prepared_1 <- dat_prepared %>%
dplyr::filter(position_end_of_last_period != 'Not Invested') %>%
group_by(position_end_of_last_period, price_move_from_last_corrected) %>%
summarize(mean_update = mean(belief_diff_from_last),
          SE = sd(belief_diff_from_last) / sqrt(n()),
          CI_90_lower = mean_update - SE * qt(.95, n()),
          CI_90_upper = mean_update + SE * qt(.95, n()))

dat_prepared_2 <- dat_prepared %>%
  dplyr::filter(position_end_of_last_period == 'Not Invested') %>%
  summarize(mean_update = mean(
              belief_diff_from_last, na.rm = TRUE),
            SE = sd(belief_diff_from_last, na.rm = TRUE) /
              sqrt(length(belief_diff_from_last)),
            CI_90_lower = mean_update - SE * qt(.95, n()),
            CI_90_upper = mean_update + SE * qt(.95, n())) %>%
  add_column(tibble(position_end_of_last_period = 'Not Invested',
    price_move_from_last_corrected = ''))

dat_prepared <- full_join(dat_prepared_1, dat_prepared_2)

ggplot(dat_prepared,
       aes(x = position_end_of_last_period, y = mean_update)) +
  facet_grid(cols = vars(price_move_from_last_corrected),
    scale = 'free_x', space = 'free_x') +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(aes(ymin = CI_90_lower, ymax = CI_90_upper),
                position = position_dodge(.9),
                width = .1) +
  labs(x = 'Price Movement and Position',
    y = 'Mean Corrected Belief Update', fill = 'Treatment')

# ggsave('updating_full_info_bars.pdf',  device = 'pdf',
#   width = 10, height = 7, path = file.path('Output', 'Plots', 'Bars'))


```

```{r update_states_shown}
dat_prepared <- dat_main_long %>%
  dplyr::filter(position_end_of_last_period != 'No Returns',
         condition == 'states_shown',
         i_block > 1) %>%
  mutate(belief_diff_bayes_corrected =
    belief_diff_from_last / 100)

dat_prepared_1 <- dat_prepared %>%
filter(position_end_of_last_period != 'Not Invested') %>%
group_by(position_end_of_last_period, price_move_from_last_corrected) %>%
summarize(mean_update = mean(belief_diff_bayes_corrected),
          SE = sd(belief_diff_bayes_corrected) / sqrt(n()),
          CI_90_lower = mean_update - SE * qt(.95, n()),
          CI_90_upper = mean_update + SE * qt(.95, n()))

dat_prepared_2 <- dat_prepared %>%
  filter(position_end_of_last_period == 'Not Invested') %>%
  summarize(mean_update = mean(
              belief_diff_bayes_corrected, na.rm = TRUE),
            SE = sd(belief_diff_bayes_corrected, na.rm = TRUE) / sqrt(n()),
            CI_90_lower = mean_update - SE * qt(.95, n()),
            CI_90_upper = mean_update + SE * qt(.95, n())) %>%
  add_column(tibble(position_end_of_last_period = 'Not Invested',
    price_move_from_last_corrected = ''))

dat_prepared <- full_join(dat_prepared_1, dat_prepared_2) %>%
  add_column(width = c(1, 1, 1, 1, .5) * .95)

ggplot(dat_prepared,
       aes(x =  fct_relevel(position_end_of_last_period,
          'Not Invested', 'Gain', 'Loss'),
          y = mean_update, fill = price_move_from_last_corrected,
          width = width)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  geom_errorbar(aes(ymin = CI_90_lower, ymax = CI_90_upper),
                    width = .1, color = 'black',
                    position = position_dodge(1)) +
  scale_fill_manual(name = 'Price Move', values = this_col_set[1:3],
    breaks = c('Favorable', 'Unfavorable')) +
  labs(x = 'Position',
    y = 'Average Corrected Belief Update') +
  theme(text = element_text(size = 16))

ggsave('updating_full_info_bars.pdf',  device = 'pdf',
  width = 10, height = 7, path = file.path('Output', 'Plots', 'Bars'))
```